---
title: "MNIST Recognition"
output:
  html_document:
    df_print: paged
---

### Part 1: Data Work

>Read both the training data and testing data.

```{r}
# read in data
library(readr)

#read training data
train_orig <- read_csv("train.csv")

#read testing data
test_orig <- read_csv("test.csv")
```

>Split the label from the pixels and transform it to a category (factor)

```{r}
# save the training labels
train_orig_labels <- train_orig[, 1]

#convert it to factor (for classification)
train_orig_labels <- as.factor(train_orig_labels$label)

#this shows how many instances of each number we have
summary(train_orig_labels)
```

>In the summary, we can see that there are about the same number of instances for each category.

### Part 2: Training the algorithm.

We will be using a `random forest` to predict the number entered.

>training the algorithm (takes about 3 mins).

```{r}
library(randomForest)
numTrees <- 25

# Train on entire training dataset and predict on the test
startTime <- proc.time()
rf <- randomForest(train_orig[-1], train_orig_labels, xtest=test_orig, ntree=numTrees, keep.forest=TRUE)
proc.time() - startTime
```

>summarizing the random forest results:

```{r}
rf
```

> Plotting the RF object.

```{r}
plot(rf, type= "l")
```

> Here we are looking at a row (a set of pixels) from the testing data, and plotting it. 

```{r}
#install.packages("RSEIS")
library(RSEIS)

#pick an image from the test poll
number <- train_orig[12, 2:785]

#flip matrix
m = matrix(number, nrow = 28, ncol = 28, byrow = FALSE)
im_numbers <- apply(m, 2, as.numeric)
im_numbers <- mirror.matrix(im_numbers)

#show image
image(1:28, 1:28, im_numbers, col=gray((0:255)/255))
```

> finally we can predict what number it is: 

```{r}
#make prediction
pred <- predict(object = rf, newdata = number, type= "response")

#shwo prediction
as.numeric(pred) - 1

```

